---
title: "PSTAT 131 HW 3"
author: "Evan Jamison"
date: "2025-11-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ISLR)
library(glmnet)
library(tree)
library(maptree)
library(randomForest)
library(gbm)
library(ROCR)
```

Predicting carseats sales using regularized regression methods

```{r}
set.seed(123)
dat <- model.matrix(Sales~., Carseats)
train = sample(nrow(dat), 50)
x.train = dat[train, ]
y.train = Carseats[train, ]$Sales
# The rest as test data
x.test = dat[-train, ]
y.test = Carseats[-train, ]$Sales
```

a.
```{r}
lambda.list.ridge = 1000 * exp(seq(0, log(1e-5), length = 100))
ridge = cv.glmnet(
  x.train, 
  y.train,
  alpha = 0,      
  lambda = lambda.list.ridge,
  nfolds = 5
)

blr = ridge$lambda.min
blr

ridge_final = glmnet(
  x.train,
  y.train,
  alpha = 0,
  lambda = blr
)

coef(ridge_final)
```

b.
```{r}
ridge_train_pred = predict(ridge_final, newx = x.train)
training_mse = mean((ridge_train_pred - y.train)^2)
training_mse

ridge_test_pred = predict(ridge_final, newx = x.test)
test_mse = mean((ridge_test_pred - y.test)^2)
test_mse

```
I find that the test MSE is significantly higher than the training MSE. However, this is common/normal behavior. 

c.
```{r}
lambda.list.lasso = 2 * exp(seq(0, log(1e-4), length = 100))

lasso = cv.glmnet(
  x.train,
  y.train,
  alpha = 1,               
  lambda = lambda.list.lasso,
  nfolds = 10
)

bll = lasso$lambda.min
bll

lasso_final = glmnet(
  x.train,
  y.train,
  alpha = 1,
  lambda = bll
)

coef(lasso_final)


```
the lasso model did shrink many coefficients towards 0 (some very close) but only USYes to 0. This means the lasso penalty was not strong enough to completely neutralize the predictors.


d.
```{r}
lasso_train_pred = predict(lasso_final, newx = x.train)
lasso_train_mse = mean((lasso_train_pred - y.train)^2)
lasso_train_mse

lasso_test_pred = predict(lasso_final, newx = x.test)
lasso_test_mse = mean((lasso_test_pred - y.test)^2)
lasso_test_mse

```
I find that similarly to ridge regression, the test MSE is significantly higher than the training MSE.


e.

Both Ridge and Lasso in this instance behaved quite similarly, both producing very similar training and test mses, with ridge performing just slightly better. Both lambda values chosen by each model were very small, however the lasso model produced a smaller lambda. Generally, ridge only shrinks coeficients while lasso while often completely set some to zero, but perhaps since the trianing size was only 50, the lasso model did not get rid of any predictors so the two models performed very similarly. 



Using bootstrap to compute uncertainty about a parameter of interes

```{r}
shots  <- 784
makes  <- 311
B <- 10000  
shots_total <- c(rep(1, makes), rep(0, shots - makes))

bootstrap_fg <- replicate(B, {
  mean(sample(shots, size = shots, replace = TRUE))
})

ci_99 <- quantile(bootstrap_fg, prob = c(0.005, 0.995))
ci_99
```



Analyzing drug use

a.
```{r}
setwd("C:/Users/user/OneDrive/Documents")

drug <- read_csv('drug.csv',
                col_names = c('ID','Age','Gender','Education','Country',
                'Ethnicity','Nscore','Escore','Oscore','Ascore','Cscore',
                'Impulsive','SS','Alcohol','Amphet','Amyl','Benzos',
                'Caff','Cannabis','Choc','Coke','Crack','Ecstasy',
                'Heroin','Ketamine','LegalH','LSD','Meth','Mushrooms',
                'Nicotine','Semer','VSA')
)

drug <- drug %>%
  mutate(
    recent_nicotine_use = ifelse(Nicotine >= "CL3", "Yes", "No"),
    recent_nicotine_use = factor(recent_nicotine_use)
  )
head(drug)
```

b.
```{r}
drug_sub <- drug %>%
  select(Age:SS, recent_nicotine_use)
head(drug_sub)
```

c.
```{r}
train_idx <- sample(nrow(drug_sub), 1000)
train <- drug_sub[train_idx, ]
test  <- drug_sub[-train_idx, ]
```

d.
```{r}
logit_reg <- glm(recent_nicotine_use ~ ., 
                 data = train, 
                 family = "binomial")
summary(logit_reg)
```

e.
```{r}
tree_model <- tree(recent_nicotine_use ~ ., data = train)
summary(tree_model)
plot(tree_model)
text(tree_model, pretty = 0)
```

f.
```{r}
cv_model <- cv.tree(tree_model, FUN = prune.misclass)
cv_model$size   
cv_model$dev    
best_s <- cv_model$size[which.min(cv_model$dev)]
best_s
```

g.
```{r}
pruned <- prune.misclass(tree_model, best = best_s)
draw.tree(pruned, nodeinfo = TRUE)
```
The variable that is split first is Age.

h.
```{r}
tree_pred <- predict(pruned, newdata = test, type = "class")
confusion_matrix <- table(True = test$recent_nicotine_use,
                  Pred = tree_pred)
confusion_matrix
TP  <- confusion_matrix["Yes", "Yes"]
FN  <- confusion_matrix["Yes", "No"]
FP  <- confusion_matrix["No",  "Yes"]
TN  <- confusion_matrix["No",  "No"]
TPR <- TP / (TP + FN)
FPR <- FP / (FP + TN)
TPR
FPR
```


i.
```{r}
train$recent_nicotine_use_num <- ifelse(train$recent_nicotine_use == "Yes", 1, 0)
test$recent_nicotine_use_num  <- ifelse(test$recent_nicotine_use  == "Yes", 1, 0)
boost_model <- gbm(
  recent_nicotine_use_num ~ . - recent_nicotine_use,  
  data = train,
  distribution = "bernoulli",
  n.trees = 1000,
  shrinkage = 0.01
)
summary(boost_model)
```
Age seems to be the most important predictor

j.
```{r}
rf <- randomForest(recent_nicotine_use ~ .,
                       data = train,
                       importance = TRUE)
rf
rf$err.rate[nrow(rf$err.rate), "OOB"]
rf$mtry
rf$ntree
varImpPlot(rf)
```
Yes, the order of importance is similar, but they're not the same.

k.
```{r}
boost_prob <- predict(boost_model, newdata = test, n.trees = 1000, type = "response")
boost_pred <- ifelse(boost_prob >= 0.20, "Yes", "No")
boost_pred <- factor(boost_pred, levels = c("No", "Yes"))
rf_prob <- predict(rf, newdata = test, type = "prob")[, "Yes"]
rf_pred <- ifelse(rf_prob >= 0.20, "Yes", "No")
rf_pred <- factor(rf_pred, levels = c("No", "Yes"))
boost_conf <- table(True = test$recent_nicotine_use, Pred = boost_pred)
rf_conf    <- table(True = test$recent_nicotine_use, Pred = rf_pred)
boost_conf
rf_conf
sum(test$recent_nicotine_use[rf_pred == "Yes"] == "Yes") / sum(rf_pred == "Yes")
```